### 2019-01-10 papers/paper reviews (for ethics course)/rec_2.tex

\documentclass{article}
% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{geometry}
    \geometry{
        a4paper,
        total={170mm,257mm},
        left=20mm,
        top=10mm,
        }

\begin{document}

\title{Peer single blind review of "Control loops in automotive systems" paper}

\author{}

\maketitle

\section{Plagiarism assesment}

The tool which was used: www.quetext.com Results: no plagiarism detected.

\section{Research specifics}

\begin{itemize}

    \item State clearly the novelty of the control loops used
    \item State clearly the novelty of some of the tools and techniques used for verifying the control loops: Unity, UDP, unit tests, systems testing.

\end{itemize}

\section{Results specifics}

\begin{itemize}

    \item Provide data interpretation of the control loop results, using their time diagrams
    \item Discuss the performance of each control loop
    \item Add and briefly discuss future improvements which can be made to this system

\end{itemize}

\section{Recommended decision}

Minor revisions are required for this paper.

\end{document}