### 2020-10-11 thesis/report2.tex
Initial checkin for report number 2.

\documentclass[a4paper,12pt,twoside]{book}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{apacite}
\usepackage{natbib}
\usepackage{listings}
\usepackage{graphicx}

\bibliographystyle{apacite}
\pagestyle{myheadings}

\title{ City traffic, IoT and Blockchain technologies }
\author{Toma Becea}

\begin{document}

\maketitle

\clearpage

\section{Introduction}

\section{Abstract}

\section{}
\label{sec:microservicesbasedapproachtrafficsim}

\section{}
\label{sec:citytrafficoptimizations}

\bibliography{wiki}

\end{document}